A Simplified Plackett-Luce Likelihood Function for Rank Orders

### 2025-10-22 inst/exponential_BT.Rmd
slight reworking

```{r fourcompetitors}
worker <- function(...){
PL <- function(v){
  stopifnot(length(v) == 4)
  v[1]/(v[1]+v[2]+v[3]+v[4]) *
  v[2]/(     v[2]+v[3]+v[4]) *
  v[3]/(          v[3]+v[4])
}

@@

@@

abline(h=0,lty=2)
abline(h = -2, lty=1)

legend("topright",lty=1,col=c("black","red","blue","green"),
       legend=c("r=1","r=2","r=3","r=4"), bg="white")
}

worker()
pdf(file="1234third.pdf")
worker()
dev.off()
```

>>>>

```{r fourcompetitors}
PL <- function(v){
  stopifnot(length(v) == 4)
  v[1]/(v[1]+v[2]+v[3]+v[4]) *
  v[2]/(     v[2]+v[3]+v[4]) *
  v[3]/(          v[3]+v[4])
}

@@

worker <- function(...){

@@

abline(h=0,lty=2)
abline(h = -2, lty=1, col="gray")

legend("topright",lty=1,col=c("black", "red", "blue", "green"),
       legend=c("r=1", "r=2", "r=3", "r=4"), bg="white")

f1_solve <- function(x){log(sapply(exp(x), f1)) + 2}
f2_solve <- function(x){log(sapply(exp(x), f2)) + 2}
f4_solve <- function(x){log(sapply(exp(x), f4)) + 2}

root1 <- uniroot(f1_solve, c(-3,-0.1))$root
root2 <- uniroot(f2_solve, c(-3,-0.1))$root
root4 <- uniroot(f4_solve, c(-3,-0.1))$root

segments(x0=root1, y0=-2.1, y1=-1.9, col="gray")
segments(x0=root2, y0=-2.1, y1=-1.9, col="gray")
segments(x0=root4, y0=-2.1, y1=-1.9, col="gray")

print(c(root1, root2, root4))
} # worker() def closes

worker()
pdf(file="1234third.pdf")
worker()
dev.off()

# Try 7 competitors, focal competitor comes third

```{r}

library(partitions)
perms(7) -> x
## first r=1

jj <- x[,x[3,] == 1]

### 2025-10-22 inst/hankin-exponential_BT.tex
slight reworking

\begin{eqnarray}
\mathcal{L}(x;r=1)
\quad &=& \quad\hphantom{+}
            \plackett{2}{3}{1}{4}\\
&{}&\quad + \plackett{2}{4}{1}{3}\\
&{}&\quad + \plackett{3}{2}{1}{4}\\
&{}&\quad + \plackett{3}{4}{1}{2}\\
&{}&\quad + \plackett{4}{2}{1}{3}\\
&{}&\quad + \plackett{4}{3}{1}{2}
\end{eqnarray}

@@

\begin{figure}[t]
  \begin{centering}
\includegraphics[width=4in]{1234third}  % 1234third.pdf produced by
                                       % very_simplified_likelihood.Rmd
\caption{Likelihood function for the observation that the focal
  competitor came third out of four, showing
  $r=1,2,3,4$\label{1234third}.  Value of $r$ indicated in colour by
  its curve; dotted line shows (asymptotic) maximum likelihood
  estimate (viz, $r=3, x\longrightarrow 0$)}
\end{centering}
\end{figure}

>>>>

\begin{eqnarray}
\mathcal{L}(x;r=1)
\quad &=& \quad\hphantom{+}
            \plackett{2}{3}{1}{4}\nonumber\\
&{}&\quad + \plackett{2}{4}{1}{3}\nonumber\\
&{}&\quad + \plackett{3}{2}{1}{4}\nonumber\\
&{}&\quad + \plackett{3}{4}{1}{2}\nonumber\\
&{}&\quad + \plackett{4}{2}{1}{3}\nonumber\\
&{}&\quad + \plackett{4}{3}{1}{2}
\end{eqnarray}

@@

\begin{figure}[t]
  \begin{centering}
\includegraphics[width=4in]{1234third}  % 1234third.pdf produced by
                                        % exponential_BT.Rmd
\caption{Support (log-likelihood) function for the observation that
  the focal competitor came third out of four, showing
  $r=1,2,3,4$\label{1234third}.  Value of $r$ indicated in colour by
  its curve; dotted line shows (asymptotic) maximum likelihood
  estimate, viz $r=3, x\longrightarrow 0$.  Gray line at $y=-2$ shows
  the two-units-of-support limit; solutions of $\mathcal{S}=-2$
  indicated by short vertical gray lines.  For example, looking at the
  red line, we may reject the hypothesis $r=2,
  x\leqslant\exp(-1.93)\simeq 0.147$ (in favour of $r=3, x=0^+$).
  Observe that all four lines meet at $x=1$, at which point all four
  competitors have identical Plackett-Luce strengths and this
  observation is uninformative as to the value of $r$}
\end{centering}
\end{figure}
